---
title: "2016-0509 MSDS 6304-401 Case Study 2"
author: "Nathan Mowat Chris Woodard, Jessica Wheeler, Bill Kerneckel"
date: "July 7, 2016"
output:
  html_document:
    keep_md: yes
---

<br>

#### Introduction


Chulwalar is part of the island group Urbano in the northern hemisphere. They 
are famous for their plants which flower in winter. There are three main plants
that Chulwalar exports: Efak is a leafy bush with white flowers, Wuge is a grass 
like plant with tiny pink flowers and Etel is a flowering tree. Etel comes in 
two varieties: red flowers and blue flowers. Due to the nature of the products,
exports generally are higher towards the end of the year. 
Chulwalar celebrates its independence on 1st December each year. On this day it
is custom to give presents to family and friends. Chulwalar also celebrates the 
March Equinox as a time of rebirth in the northern hemisphere. 
<br>
<br>
The Prime Minister of Chulwalar has asked us to help him in forecasting the 
exports. In order to do this we have been given as is data and plan data as well
as a list of indicators which may affect exports. Our job is to find out the best
way to forecast Chulwalar's exports in 2014 based on data collected before this year.

<strong>Assignment</strong>

Choose one of the flowers from the data provided and complete the following analysis:

- Section the code into modules for importing data, cleaning/modifying data, exploring data, analyzing data

- Add interpretation

- Determine which model is best

- Give a conclusion of your analysis as far as forecasting exports for Chulwalar

- For this project, we will be forecasting the "<font color="red">RedEtel</font>" flower exports for Chulwalar. 

****************************

#### Table of Contents

* 1.0   [Preperation, import and convert data](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/1.md)
* 1.1   [Import the exports data and the indicators](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/1.md)
* 1.2   [Transformation the data into vectors and time series](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/1.md)

* 2.0   [Analysis of the basic data](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md) 
* 2.1   [Development of the business portfolio](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md)
* 2.2   [Correlation between As Is and Plan data](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md)
* 2.3   [Time series analysis](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md)
* 2.3.1 ["stl" function](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md)
* 2.3.2 [Modification of the seasonal componant to a monthly base.](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/2.md)

* 3.0 [Correlation of different external indicators](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/3.md)
* 3.1 [Definition of the indicators and their correlation with the basic data](https://github.com/wkerneck/CaseStudy2/blob/master/analysis/data/3.md)
  
* 19.0  [Conclusion and Summary.](#id-section19) 
* 20.0  [Acknowledgements](#id-section20) 

****************************


#### Preliminary Steps

In order for the analysis of the three datasets you must set your working directory to the following:

```{r}
setwd("/Users/wkerneck/desktop/CaseStudy2/")
```

The libraries listed below must be installed in order for the functions outlined in the steps below to execute. Load 'fpp' package in order to obtain the forecasting functions. Load 'tcltk' for pause function.

```{r message=FALSE, warning=FALSE}
library(fpp)
library(tcltk)
mywait <- function() {
    tt <- tktoplevel()
    tkpack( tkbutton(tt, text='Continue', command=function()tkdestroy(tt)),
        side='bottom')
    tkbind(tt,'<Key>', function()tkdestroy(tt) )

    tkwait.window(tt)
}

cat("Success: All libaries downloaded")
```


