---
title: "2016-0509 MSDS 6304-401 Case Study 2"
author: "Claire Chu, Chris Woodard, Jessica Wheeler, Bill Kerneckel"
date: "July 7, 2016"
output:
  html_document:
    keep_md: yes
---

<br>

#### Introduction


Chulwalar is part of the island group Urbano in the northern hemisphere. They 
are famous for their plants which flower in winter. There are three main plants
that Chulwalar exports: Efak is a leafy bush with white flowers, Wuge is a grass 
like plant with tiny pink flowers and Etel is a flowering tree. Etel comes in 
two varieties: red flowers and blue flowers. Due to the nature of the products,
exports generally are higher towards the end of the year. 
Chulwalar celebrates its independence on 1st December each year. On this day it
is custom to give presents to family and friends. Chulwalar also celebrates the 
March Equinox as a time of rebirth in the northern hemisphere. 
<br>
<br>
The Prime Minister of Chulwalar has asked us to help him in forecasting the 
exports. In order to do this we have been given as is data and plan data as well
as a list of indicators which may affect exports. Our job is to find out the best
way to forecast Chulwalar's exports in 2014 based on data collected before this year
- thus to make any statistical model we introduce credible. 

****************************

#### Setting your working directory

In order for the analysis of thedatasets you must set your working directory to the following:

```{r}
setwd("/Users/wkerneck/desktop/CaseStudy2/")
```

****************************

#### Table of Contents

- 1.0   [Preperation, import and convert data](#id-section1)
- 1.1   [Import the exports data and the indicators](#id-section1.1)
- 1.2   [Transformation the data into vectors and time series.](#id-section1.2)

- 2.0   [Analysis of the basic data.](#id-section2) 
- 2.1   [Development of the business portfolio.](#id-section2.1)
- 2.2   [Correlation between As Is and Plan data.](#id-section2.2)
- 2.3   [Time series analysis.](#id-section2.3)
- 2.3.1 ["stl" function.](#id-section2.3.1)
- 2.3.2 [Modification of the seasonal componant to a monthly base.](#id-section2.3.2)

- 3.0 [Correlation of different external indicators .](#id-section3)
- 3.1 [Definition of the indicators and their correlation with the basic data.](#id-section3.1)
  
- 19.0  [Conclusion and Summary](#id-section19) 
- 20.0  [Acknowledgements](#id-section20) 

****************************
<div id='id-section1'/>
####  1. Preperation, import and convert data

The libraries listed below must be installed in order for the functions outlined in the steps below to execute. Load 'fpp' package in order to obtain the forecasting functions. Load 'tcltk' for pause function.

```{r message=FALSE, warning=FALSE}
library(fpp)
library(tcltk)
mywait <- function() {
    tt <- tktoplevel()
    tkpack( tkbutton(tt, text='Continue', command=function()tkdestroy(tt)),
        side='bottom')
    tkbind(tt,'<Key>', function()tkdestroy(tt) )

    tkwait.window(tt)
}

cat("Success: All libaries downloaded")
```



****************************
<div id='id-section1.1'/>
####  1.1 Import the exports data and the indicators

In order to test the script, it is necessary to change the three file paths. The files have been sent together with the script.

- The Export data for Chulwalar   are in two .csv files.
- One file for the as is data: ImportedAsIsDataChulwalar.csv
- and another one for the plan data: ImportedPlanDataChulwalar.csv

```{r}
ImportedAsIsData <- read.csv("ImportedAsIsDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedPlanData <- read.csv("ImportedPlanDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedIndicators <- read.csv("ImportedIndicatorsChulwalar.csv", header = F, sep=";", fill = T)

head(ImportedAsIsData)
head(ImportedPlanData)
head(ImportedIndicators)
```

****************************
<div id='id-section1.2'/>
####  1.2 Transformation the data into vectors and time series.

In order to be able to work with the partial data sets later, these need to be split into individual vectors and converted into times series.

```{r}
TotalAsIsVector <- c(ImportedAsIsData [2:13,2],ImportedAsIsData [2:13,3],ImportedAsIsData [2:13,4],ImportedAsIsData [2:13,5],ImportedAsIsData [2:13,6],ImportedAsIsData [2:13,7])

EfakAsIsVector <- c(ImportedAsIsData [16:27,2],ImportedAsIsData [16:27,3],ImportedAsIsData [16:27,4],ImportedAsIsData [16:27,5],ImportedAsIsData [16:27,6],ImportedAsIsData [16:27,7])

WugeAsIsVector <- c(ImportedAsIsData [30:41,2],ImportedAsIsData [30:41,3],ImportedAsIsData [30:41,4],ImportedAsIsData [30:41,5],ImportedAsIsData [30:41,6],ImportedAsIsData [30:41,7])

TotalEtelAsIsVector <- c(ImportedAsIsData [44:55,2],ImportedAsIsData [44:55,3],ImportedAsIsData [44:55,4],ImportedAsIsData [44:55,5],ImportedAsIsData [44:55,6],ImportedAsIsData [44:55,7])

BlueEtelAsIsVector <- c(ImportedAsIsData [58:69,2],ImportedAsIsData [58:69,3],ImportedAsIsData [58:69,4],ImportedAsIsData [58:69,5],ImportedAsIsData [58:69,6],ImportedAsIsData [58:69,7])

RedEtelAsIsVector <- c(ImportedAsIsData [72:83,2],ImportedAsIsData [72:83,3],ImportedAsIsData [72:83,4],ImportedAsIsData [72:83,5],ImportedAsIsData [72:83,6],ImportedAsIsData [72:83,7])

YearAsIsVector <- c(ImportedAsIsData [86,2],ImportedAsIsData [86,3],ImportedAsIsData [86,4],ImportedAsIsData [86,5],ImportedAsIsData [86,6],ImportedAsIsData [86,7])

TotalAsIsVector_2014 <- c(ImportedAsIsData[2:13,8])

PlanVector <- c(ImportedPlanData[2:13,2],ImportedPlanData[2:13,3],ImportedPlanData[2:13,4],ImportedPlanData[2:13,5],ImportedPlanData[2:13,6],ImportedPlanData[2:13,7])

EfakPlanVector <- c(ImportedPlanData[16:27,2],ImportedPlanData[16:27,3],ImportedPlanData[16:27,4],ImportedPlanData[16:27,5],ImportedPlanData[16:27,6],ImportedPlanData[16:27,7])

WugePlanVector <- c(ImportedPlanData[30:41,2],ImportedPlanData[30:41,3],ImportedPlanData[30:41,4],ImportedPlanData[30:41,5],ImportedPlanData[30:41,6],ImportedPlanData[30:41,7])

TotalEtelPlanVector <- c(ImportedPlanData[44:55,2],ImportedPlanData[44:55,3],ImportedPlanData[44:55,4],ImportedPlanData[44:55,5],ImportedPlanData[44:55,6],ImportedPlanData[44:55,7])

BlueEtelPlanVector <- c(ImportedPlanData[58:69,2],ImportedPlanData[58:69,3],ImportedPlanData[58:69,4],ImportedPlanData[58:69,5],ImportedPlanData[58:69,6],ImportedPlanData[58:69,7])

RedEtelPlanVector <- c(ImportedPlanData[72:83,2],ImportedPlanData[72:83,3],ImportedPlanData[72:83,4],ImportedPlanData[72:83,5],ImportedPlanData[72:83,6],ImportedPlanData[72:83,7])

YearPlanVector <- c(ImportedPlanData[86,2],ImportedPlanData[86,3],ImportedPlanData[86,4],ImportedPlanData[86,5],ImportedPlanData[86,6],ImportedPlanData[86,7])
PlanVector_2014 <- c(ImportedPlanData[2:13,8])
```

The data is saved as a vector and needs to be converted into a time series.

```{r}
TotalAsIs<- ts(TotalAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakAsIs <- ts(EfakAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
WugeAsIs <- ts(WugeAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelAsIs<- ts(TotalEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelAsIs <- ts(BlueEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelAsIs <- ts(RedEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearAsIs <- ts(YearAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalAsIs_2014 <- ts(TotalAsIsVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)

TotalPlan <- ts(PlanVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakPlan <- ts(EfakPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
WugePlan <- ts(WugePlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelPlan <- ts(TotalEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelPlan <- ts(BlueEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelPlan <- ts(RedEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearPlan <- ts(YearPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalPlan_2014 <- ts(PlanVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)
```

Call up the time series to check everything has worked.

```{R}
TotalAsIs
EfakAsIs
WugeAsIs
TotalEtelAsIs
BlueEtelAsIs 
RedEtelAsIs 
YearAsIs
TotalAsIs_2014

TotalPlan
EfakPlan
WugePlan 
TotalEtelPlan
BlueEtelPlan
RedEtelPlan
YearPlan
TotalPlan_2014
```
****************************
<div id='id-section2.0'/>
#### 2. Analysis of the basic data  
<div id='id-section2.1'/>
#### 2.1 Development of the business portfolio

Due to the different scales, it makes sense to plot each graph individually instead of plotting them all on one set of axes. 

```{r}
par(mfrow=c(3,2))

plot(TotalAsIs, col="black", main="TotalAsIs")
plot(EfakAsIs , col="red",main="EfakAsIs")
plot(WugeAsIs, col="blue", main="WugeAsIs")
plot(TotalEtelAsIs, col="green",main="TotalEtelAsIs")
plot(BlueEtelAsIs, col="orange", main="BlueEtelAsIs")
plot(RedEtelAsIs, col="purple", main="RedEtelAsIs")

mywait()
mywait()

plot(TotalPlan , col="black", main="TotalPlan")
plot(EfakPlan , col="red",main="EfakPlan")
plot(WugePlan, col="blue", main="WugePlan")
plot(TotalEtelPlan, col="green",main="TotalEtelPlan")
plot(BlueEtelPlan, col="orange", main="BlueEtelPlan")
plot(RedEtelPlan, col="purple", main="RedEtelPlan")

mywait()
```

****************************
<div id='id-section2.1'/>
#### 2.1 Development of the business portfolio

Test the correlation between As Is and Plan data in order to test how exact the planning is. Correlation is a measure of linear relationship between two variables. 

```{r}
cor(TotalAsIs, TotalPlan )
cor(EfakAsIs , EfakPlan)
cor(WugeAsIs, WugePlan)
cor(TotalEtelAsIs, TotalEtelPlan)
cor(BlueEtelAsIs , BlueEtelPlan)
cor(RedEtelAsIs , RedEtelPlan)
cor(YearAsIs, YearPlan)

mywait()
```

The results show a very high planning accuracy

```{r}
TotalAsIs_lm <- lm(TotalAsIs ~ TotalPlan , data = TotalAsIs)
summary(TotalAsIs_lm)

TotalAsIs_tslm <- tslm(TotalAsIs ~ TotalPlan )
summary(TotalAsIs_tslm)
```

****************************
<div id='id-section2.3'/>
#### 2.3 Time series analysis

****************************
<div id='id-section2.3.1'/>
#### 2.3.1 "stl" function

The time series can be analysed using the stl function in order to seperate the trend, seasonality and remainder (remaining coincidential) components from one another.

```{r}
TotalAsIs_stl <- stl(TotalAsIs, s.window=5)
EfakAsIs_stl <- stl(EfakAsIs , s.window=5)
WugeAsIs_stl <- stl(WugeAsIs, s.window=5)
TotalEtelAsIs_stl <- stl(TotalEtelAsIs, s.window=5)
BlueEtelAsIs_stl <- stl(BlueEtelAsIs , s.window=5)
RedEtelAsIs_stl <- stl(RedEtelAsIs , s.window=5)
```

Thus the individual time series can be shown graphically and tabularly. The trend of the total exports is almost linear. A relatively uniform seaonality can be seen.

```{r}
par(mfrow=c(3,2))

plot(TotalAsIs_stl, col="black", main="TotalAsIs_stl")
TotalAsIs_stl
```

It is interesting to note that the almost linear trend is not seen in the individual segments. The individual trends run partially in opposite directions in the middle of the time scale, which causes the linear trend in the total As Is data.

```{r}
plot(EfakAsIs_stl, col="black", main="EfakAsIs_stl")
EfakAsIs_stl

plot(WugeAsIs_stl, col="black", main="WugeAsIs_stl")
WugeAsIs_stl

plot(TotalEtelAsIs_stl, col="black", main="TotalEtelAsIs_stl")
TotalEtelAsIs_stl

plot(BlueEtelAsIs_stl, col="black", main="BlueEtelAsIs_stl")
BlueEtelAsIs_stl

plot(RedEtelAsIs_stl, col="black", main="RedEtelAsIs_stl")
RedEtelAsIs_stl

mywait()

par(mfrow=c(3,2))

plot(TotalAsIs_stl$time.series[,"trend"], col="black")
plot(EfakAsIs_stl$time.series[,"trend"], col="red")
plot(WugeAsIs_stl$time.series[,"trend"], col="blue")
plot(TotalEtelAsIs_stl$time.series[,"trend"], col="green")
plot(BlueEtelAsIs_stl$time.series[,"trend"], col="orange")
plot(RedEtelAsIs_stl$time.series[,"trend"], col="purple")

mywait()
```

****************************
<div id='id-section2.3.2'/>
#### 2.3.2 Modification of the seasonal componant to a monthly base

The modification of the seasonlity component can also be changed into a monthly view. It only makes sense to do this if the seasonality componant as the trend looks almost identical and the remainder is then randomly spread. 

```{r}
monthplot(TotalAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
monthplot(EfakAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
monthplot(WugeAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
monthplot(TotalEtelAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
monthplot(BlueEtelAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")
monthplot(RedEtelAsIs_stl$time.series[,"seasonal"], main="", ylab="Seasonal")

mywait()
```

****************************
<div id='id-section3'/>
#### 3 Correlation of different external indicators

****************************
<div id='id-section3.1'/>
#### 3.1 Definition of the indicators and their correlation with the basic data

The following indicators are to be tested:

1   Monthly Change in Export Price Index (CEPI)
2 	Monthly Satisfaction Index (SI) government based data
3 	Average monthly temperatures in Chulwalar
4 	Monthly births in Chulwalar
5 	Monthly Satisfaction Index (SI) external index 
6 	Yearly Exports from Urbano
7 	Yearly number of Globalisation Party members in Chulwalar
8 	Monthly Average Export Price Index for Chulwalar
9 	Monthly Producer Price Index (PPI) for Etel in Chulwalar
10 	National Holidays
11 	Chulwalar Index (Total value of all companies in Chulwalar)
12	Monthly Inflation rate in Chulwalar
13	Proposed spending for National Holidays
14    Influence of National Holiday

The indicators will be converted into individual  vectors and subsequently converted into time series. The correlation of the indicators will then be tested against the As Is exports for Chulwalar. 

Monthly Change in Export Price Index (CEPI)

```{r}
CEPIVector <- c(ImportedIndicators[2:13,2],ImportedIndicators[2:13,3],ImportedIndicators[2:13,4],ImportedIndicators[2:13,5],ImportedIndicators[2:13,6],ImportedIndicators[2:13,7])
CEPI <- ts(CEPIVector , start=c(2008,1), end=c(2013,12), frequency=12)
plot(CEPI, main="CEPI")

cor(TotalAsIs, CEPI)
cor(EfakAsIs , CEPI)
cor(WugeAsIs, CEPI)
cor(TotalEtelAsIs, CEPI)
cor(BlueEtelAsIs , CEPI)
cor(RedEtelAsIs , CEPI)

mywait()
```
****************************
<div id='id-section19'/>
#### Conclusion and Summary




****************************
<div id='id-section20'/>
#### Acknowledgements

Planning and forecasting in a volatile setting: the Chulwalar case v0.8alpha        

Contributing authors:
<ul>
<li>Amy Wheeler</li>
<li>Nina Weitkamp</li>
<li>Patrick Berlekamp</li>
<li>Johannes Brauer</li>
<li>Andreas Faatz</li>
<li>Hans-Ulrich Holst</li>
</ul>
                                                                             
Designed and coded at Hochschule Osnabrück, Germany              
Contact: faatz@wi.hs-osnabrueck.de                     

Thanks to: Rob Hyndman for all the lovely forecasting libraries in R                                                                                                    