---
title: "2016-0509 MSDS 6304-401 Case Study 2"
author: "Claire Chu, Chris Woodward, Jessica Wheeler, Bill Kerneckel"
date: "July 7, 2016"
output:
  html_document:
    keep_md: yes
---

<br>

#### Introduction


Chulwalar is part of the island group Urbano in the northern hemisphere. They 
are famous for their plants which flower in winter. There are three main plants
that Chulwalar exports: Efak is a leafy bush with white flowers, Wuge is a grass 
like plant with tiny pink flowers and Etel is a flowering tree. Etel comes in 
two varieties: red flowers and blue flowers. Due to the nature of the products,
exports generally are higher towards the end of the year. 
Chulwalar celebrates its independence on 1st December each year. On this day it
is custom to give presents to family and friends. Chulwalar also celebrates the 
March Equinox as a time of rebirth in the northern hemisphere. 
<br>
<br>
The Prime Minister of Chulwalar has asked us to help him in forecasting the 
exports. In order to do this we have been given as is data and plan data as well
as a list of indicators which may affect exports. Our job is to find out the best
way to forecast Chulwalar's exports in 2014 based on data collected before this year
- thus to make any statistical model we introduce credible. 

****************************

#### Setting your working directory

In order for the analysis of thedatasets you must set your working directory to the following:

```{r}
setwd("/Users/wkerneck/desktop/CaseStudy2/")
```

****************************

####  1. Preperation, import and convert data

The libraries listed below must be installed in order for the functions outlined in the steps below to execute. Load 'fpp' package in order to obtain the forecasting functions. Load 'tcltk' for pause function.

```{r message=FALSE, warning=FALSE}
library(fpp)
library(tcltk)
mywait <- function() {
    tt <- tktoplevel()
    tkpack( tkbutton(tt, text='Continue', command=function()tkdestroy(tt)),
        side='bottom')
    tkbind(tt,'<Key>', function()tkdestroy(tt) )

    tkwait.window(tt)
}

cat("Success: All libaries downloaded")
```


****************************

####  1.1 Import the exports data and the indicators

In order to test the script, it is necessary to change the three file paths. The files have been sent together with the script.

- The Export data for Chulwalar   are in two .csv files.
- One file for the as is data: ImportedAsIsDataChulwalar.csv
- and another one for the plan data: ImportedPlanDataChulwalar.csv

```{r}
ImportedAsIsData <- read.csv("ImportedAsIsDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedPlanData <- read.csv("ImportedPlanDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedIndicators <- read.csv("ImportedIndicatorsChulwalar.csv", header = F, sep=";", fill = T)

head(ImportedAsIsData)
head(ImportedPlanData)
head(ImportedIndicators)
```

****************************

####  1.2 Transformation the data into vectors and time series.

In order to be able to work with the partial data sets later, these need to be split into individual vectors and converted into times series.

```{r}
TotalAsIsVector <- c(ImportedAsIsData [2:13,2],ImportedAsIsData [2:13,3],ImportedAsIsData [2:13,4],ImportedAsIsData [2:13,5],ImportedAsIsData [2:13,6],ImportedAsIsData [2:13,7])

EfakAsIsVector <- c(ImportedAsIsData [16:27,2],ImportedAsIsData [16:27,3],ImportedAsIsData [16:27,4],ImportedAsIsData [16:27,5],ImportedAsIsData [16:27,6],ImportedAsIsData [16:27,7])

WugeAsIsVector <- c(ImportedAsIsData [30:41,2],ImportedAsIsData [30:41,3],ImportedAsIsData [30:41,4],ImportedAsIsData [30:41,5],ImportedAsIsData [30:41,6],ImportedAsIsData [30:41,7])

TotalEtelAsIsVector <- c(ImportedAsIsData [44:55,2],ImportedAsIsData [44:55,3],ImportedAsIsData [44:55,4],ImportedAsIsData [44:55,5],ImportedAsIsData [44:55,6],ImportedAsIsData [44:55,7])

BlueEtelAsIsVector <- c(ImportedAsIsData [58:69,2],ImportedAsIsData [58:69,3],ImportedAsIsData [58:69,4],ImportedAsIsData [58:69,5],ImportedAsIsData [58:69,6],ImportedAsIsData [58:69,7])

RedEtelAsIsVector <- c(ImportedAsIsData [72:83,2],ImportedAsIsData [72:83,3],ImportedAsIsData [72:83,4],ImportedAsIsData [72:83,5],ImportedAsIsData [72:83,6],ImportedAsIsData [72:83,7])

YearAsIsVector <- c(ImportedAsIsData [86,2],ImportedAsIsData [86,3],ImportedAsIsData [86,4],ImportedAsIsData [86,5],ImportedAsIsData [86,6],ImportedAsIsData [86,7])

TotalAsIsVector_2014 <- c(ImportedAsIsData[2:13,8])

PlanVector <- c(ImportedPlanData[2:13,2],ImportedPlanData[2:13,3],ImportedPlanData[2:13,4],ImportedPlanData[2:13,5],ImportedPlanData[2:13,6],ImportedPlanData[2:13,7])

EfakPlanVector <- c(ImportedPlanData[16:27,2],ImportedPlanData[16:27,3],ImportedPlanData[16:27,4],ImportedPlanData[16:27,5],ImportedPlanData[16:27,6],ImportedPlanData[16:27,7])

WugePlanVector <- c(ImportedPlanData[30:41,2],ImportedPlanData[30:41,3],ImportedPlanData[30:41,4],ImportedPlanData[30:41,5],ImportedPlanData[30:41,6],ImportedPlanData[30:41,7])

TotalEtelPlanVector <- c(ImportedPlanData[44:55,2],ImportedPlanData[44:55,3],ImportedPlanData[44:55,4],ImportedPlanData[44:55,5],ImportedPlanData[44:55,6],ImportedPlanData[44:55,7])

BlueEtelPlanVector <- c(ImportedPlanData[58:69,2],ImportedPlanData[58:69,3],ImportedPlanData[58:69,4],ImportedPlanData[58:69,5],ImportedPlanData[58:69,6],ImportedPlanData[58:69,7])

RedEtelPlanVector <- c(ImportedPlanData[72:83,2],ImportedPlanData[72:83,3],ImportedPlanData[72:83,4],ImportedPlanData[72:83,5],ImportedPlanData[72:83,6],ImportedPlanData[72:83,7])

YearPlanVector <- c(ImportedPlanData[86,2],ImportedPlanData[86,3],ImportedPlanData[86,4],ImportedPlanData[86,5],ImportedPlanData[86,6],ImportedPlanData[86,7])
PlanVector_2014 <- c(ImportedPlanData[2:13,8])
```

The data is saved as a vector and needs to be converted into a time series.

```{r}
TotalAsIs<- ts(TotalAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakAsIs <- ts(EfakAsIsVector , start=c(2008,1), end=c(2013,12), frequency=12)
WugeAsIs <- ts(WugeAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelAsIs<- ts(TotalEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelAsIs <- ts(BlueEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelAsIs <- ts(RedEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearAsIs <- ts(YearAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalAsIs_2014 <- ts(TotalAsIsVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)

TotalPlan <- ts(PlanVector , start=c(2008,1), end=c(2013,12), frequency=12)
EfakPlan <- ts(EfakPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
WugePlan <- ts(WugePlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalEtelPlan <- ts(TotalEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
BlueEtelPlan <- ts(BlueEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelPlan <- ts(RedEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
YearPlan <- ts(YearPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)
TotalPlan_2014 <- ts(PlanVector_2014, start=c(2014,1), end=c(2014,12), frequency=12)
```

Call up the time series to check everything has worked.

```{R}
TotalAsIs
EfakAsIs
WugeAsIs
TotalEtelAsIs
BlueEtelAsIs 
RedEtelAsIs 
YearAsIs
TotalAsIs_2014

TotalPlan
EfakPlan
WugePlan 
TotalEtelPlan
BlueEtelPlan
RedEtelPlan
YearPlan
TotalPlan_2014
```
****************************
#### 2. Analysis of the basic data  

#### 2.1 Development of the business portfolio

Due to the different scales, it makes sense to plot each graph individually instead of plotting them all on one set of axes. 

```{r}
par(mfrow=c(3,2))

plot(TotalAsIs, col="black", main="TotalAsIs")
plot(EfakAsIs , col="red",main="EfakAsIs")
plot(WugeAsIs, col="blue", main="WugeAsIs")
plot(TotalEtelAsIs, col="green",main="TotalEtelAsIs")
plot(BlueEtelAsIs, col="orange", main="BlueEtelAsIs")
plot(RedEtelAsIs, col="purple", main="RedEtelAsIs")

mywait()
mywait()

plot(TotalPlan , col="black", main="TotalPlan")
plot(EfakPlan , col="red",main="EfakPlan")
plot(WugePlan, col="blue", main="WugePlan")
plot(TotalEtelPlan, col="green",main="TotalEtelPlan")
plot(BlueEtelPlan, col="orange", main="BlueEtelPlan")
plot(RedEtelPlan, col="purple", main="RedEtelPlan")

mywait()
```

****************************
#### Conclusion and Summary




****************************

#### Acknowledgements

Planning and forecasting in a volatile setting: the Chulwalar case v0.8alpha        

Contributing authors:
<ul>
<li>Amy Wheeler</li>
<li>Nina Weitkamp</li>
<li>Patrick Berlekamp</li>
<li>Johannes Brauer</li>
<li>Andreas Faatz</li>
<li>Hans-Ulrich Holst</li>
</ul>
                                                                             
Designed and coded at Hochschule Osnabrück, Germany              
Contact: faatz@wi.hs-osnabrueck.de                     

Thanks to: Rob Hyndman for all the lovely forecasting libraries in R                                                                                                    