---
title: "Preperation, import and convert data"
author: "Nathan Mowat Chris Woodard, Jessica Wheeler, Bill Kerneckel"
date: "July 7, 2016"
output:
  html_document:
    keep_md: yes
---



****************************

#### Table of Contents

[Back to Home]("https://github.com/wkerneck/CaseStudy2")

* 1.0   [Preperation, import and convert data](#id-section1)
* 1.1   [Import the exports data and the indicators](#id-section1.1)
* 1.2   [Transformation the data into vectors and time series](#id-section1.2)


```{r echo=FALSE}
setwd("/Users/wkerneck/desktop/CaseStudy2/")
```

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(fpp)
library(tcltk)
mywait <- function() {
    tt <- tktoplevel()
    tkpack( tkbutton(tt, text='Continue', command=function()tkdestroy(tt)),
        side='bottom')
    tkbind(tt,'<Key>', function()tkdestroy(tt) )

    tkwait.window(tt)
}

```


****************************

<div id='id-section1.1'/>
####  1.1 Import the exports data and the indicators

In order to test the script, it is necessary to change the three file paths. The files have been sent together with the script.

- The Export data for Chulwalar   are in two .csv files.
- One file for the as is data: ImportedAsIsDataChulwalar.csv
- and another one for the plan data: ImportedPlanDataChulwalar.csv

```{r}
ImportedAsIsData <- read.csv("ImportedAsIsDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedPlanData <- read.csv("ImportedPlanDataChulwalar.csv", header = F, sep=";", fill = T)

ImportedIndicators <- read.csv("ImportedIndicatorsChulwalar.csv", header = F, sep=";", fill = T)

head(ImportedAsIsData)
head(ImportedPlanData)
head(ImportedIndicators)
```

****************************
<div id='id-section1.2'/>
####  1.2 Transformation the data into vectors and time series.

In order to be able to work with the partial data sets later, these need to be split into individual vectors and converted into times series.

```{r}
RedEtelAsIsVector <- c(ImportedAsIsData [72:83,2],ImportedAsIsData [72:83,3],ImportedAsIsData [72:83,4],ImportedAsIsData [72:83,5],ImportedAsIsData [72:83,6],ImportedAsIsData [72:83,7])
RedEtelPlanVector <- c(ImportedPlanData[72:83,2],ImportedPlanData[72:83,3],ImportedPlanData[72:83,4],ImportedPlanData[72:83,5],ImportedPlanData[72:83,6],ImportedPlanData[72:83,7])
```

The data is saved as a vector and needs to be converted into a time series.

```{r}

RedEtelAsIs <- ts(RedEtelAsIsVector, start=c(2008,1), end=c(2013,12), frequency=12)
RedEtelPlan <- ts(RedEtelPlanVector, start=c(2008,1), end=c(2013,12), frequency=12)

```

Call up the time series to check everything has worked.

```{R}

RedEtelAsIs 
RedEtelPlan

```
****************************
<div id='id-section2.0'/>
#### 2. Analysis of the basic data  
<div id='id-section2.1'/>
#### 2.1 Development of the business portfolio

Due to the different scales, it makes sense to plot each graph individually instead of plotting them all on one set of axes. 

```{r}
par(mfrow=c(2,1))
plot(RedEtelAsIs, col="red", main="RedEtelAsIs")
plot(RedEtelPlan, col="red", main="RedEtelPlan")

mywait()
```

****************************
